\chapter{UsiXML}

In section \ref{subsection:usixml} UsiXML was referred as part of the state of the art. In this chapter this subject will be studied in more depth. The focus of this chapter will not be just UsiXML as a specification language but the whole project behind it.

UsiXML's last version was released in February 14th, 2007. In April of 2009 was founded a consortium with the objective of lowering  the total application costs and development time by adding versatile context driven capabilities to UsiXML that would bring it far beyond the state of the art, up to the achievement of its standardisation.

The UsiXML consortium started with three main goals that would guide the entire project. The first goal is \textit{the UsiXML ``$\mu7$'' concept elicitation and promotion}. The $\mu7$ concept means that it should be possible to specify  a user interface for an interactive application that should support multi-device, multi-platform, multi-user, multi-linguality/culturality, multi-organisation, multicontext, or multi-modality capabilities.

The second goal is the \textit{development of the UsiXML language and the model-driven method}. The UsiXML language should guarantee interoperability, reusability, and maintainability of interactive applications developed. For this reason UsiXML is an open XML-compliant standard User Interface Description Language (UIDL). There should be models to cover every $\mu7$ aspects. Finally, UsiXML should define a flexible methodological framework that accommodates various development paths as found in organisations and that can be tailored to their specific needs.

The third goal is to \textit{set up development tools and demonstration of the validity on applications}. There should be a suite of software tools that support the methodological framework defined in goal 2 that can be later integrated or connected to available software environments. A UsiXML service will be defined and developed once so that this service can be deployed many times, especially for all environments requiring them to reduce the total cost of development. UsiXML will provide developers with various knowledge bases containing usability and accessibility guidelines that can be semi-automatically verified on any UsiXML-produced UI so as to guarantee a certain level of quality.

Section \ref{section:usixml_semantics} provides an in depth study of the UsiXML semantics. In section \ref{section:usixml_method} refers to the UsiXML engineering method for developing user interfaces.

\input{./content/usixml/uidl}
\input{./content/usixml/method}