\section{Specification languages}
\label{section:specification_languages}

In \cite{idealxml2} was stated that one of the most important challenges to overcome in model driven development of user interfaces is the creation of a specification language that would be massively adopted and became a common ground between developers like UML is for software architecture.

Over the years many languages were developed to try and overcome this obstacle. In \cite{mecano} were enumerated some of the problems found with that time user interface models:
\begin{itemize}
\item \textbf{Partial models}, most models deal only with a portion of the spectrum of interface characteristics. Some emphasize domain, others emphasize tasks, some others emphasize presentation guidelines and so on.

\item \textbf{Insufficient underlying model}, several model-based systems use modelling paradigms proven successful in other applications areas, but that come up short for interface development. These underlying models typically result in partial interface models of restricted expressiveness.

\item \textbf{System-dependent models}, many interface models are non-declarative and are embedded implicitly into their associated model-based systems, sometimes at code level. These generic models are tied to the interface generation schema of their system, and are therefore unusable in any other environment.

\item \textbf{Inflexible models}, experience with model-based systems suggests that interface developers often wish to change, modify, or expand the interface model associated with a particular model-based environment. However, model-based systems do not offer facilities to such modifications, nor the interface models in question are defined in a way that modifications can be easily accomplished. Thus the inclusion of an open meta-model like in UML could be an important factor of success.
\end{itemize}

Next, two recent specification languages for user interfaces that try to overcome these problems will be presented. In section \ref{subsection:umli} will be presented UMLi, an extension to UML to support the modelling of user interfaces. In section \ref{subsection:usixml} will be presented UsiXML, a language with potential to become a standard in user interface specification.

\input{content/state_of_the_art/specification_languages/umli}
\input{content/state_of_the_art/specification_languages/usixml}