\section{How user interfaces are built}
There are several techniques and several tools to build user interfaces. Some are more intuitive and easy to learn while others are more flexible but harder to learn and thus more time consuming.

In this chapter It'll be focusing on some of the most used techniques for building user interfaces. I'll try to explain what are the main advantages and disadvantages related to each technique while using some examples to justify them.
Probably the most used technique is also the oldest one, manually coding interfaces. It's hard and time consuming but it's usually preferred by most experienced developers because it's more flexible and if they're good at what they're doing the final code can be very good and maintainable.

The second technique we'll see in this chapter is code generation through WYSIWYG\footnote{What you see is what you get.} tools. There are many tools of this kind that support the most popular languages and frameworks for developing user interfaces. They're very used mainly by novice developers and designers. The final code isn't always the best but if you're using a robust tool there's little chance of finding bugs in it.

The third and last technique I'll talk about isn't the most popular in the industry but there's a lot of work surrounding it in the academic world. Model driven development is widely used for the bottom layers in software development but is not that popular for the presentation layer. Although this technique is not as widely used as the previous ones it brings many advantages such as platform independence.

\input{content/how_user_interfaces_are_built/manually_coding_user_interfaces}
\input{content/how_user_interfaces_are_built/code_generation_through_WYSIWYG_tools}
\input{content/how_user_interfaces_are_built/model_driven_development_of_user_interfaces}